---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
library(asaur)
library(survival)
library(ggplot2)
library(muhaz)
library(parfm)
data = read.csv("/Users/apple/Desktop/arrhythmia.csv")
```

```{r}
#Process data
sum(is.na(data))

data$Sex[data$Sex=="M"]=1
data$Sex[data$Sex=="F"]=0 #Male=1, Female=0
data$Race=as.numeric(as.factor(data$Race)) #Afri Amer=1, Amer Indian=2, Asian=3, No answer=4, Other=5, white=6

data$Diagnosis[data$Diagnosis=="Idiopathic"]=0
data$Diagnosis[data$Diagnosis=="Ischemic"]=1
data[,c(6:9,13:19)]=lapply(data[,c(6:9,13:19)], function(x) as.numeric(x=="Yes")) #"Yes"=1, "No"=0


abnormal=data[data$tShock>data$tDeath,]#Detect the abnormal obs.
#View(abnormal)
data=data[data$tShock<=data$tDeath,]
table(data$Race)
group1=subset(data,data$dShock==1&data$dDeath==1&(data$tDeath-data$tShock)<=1/30) # the patients died within one day after shock.
#View(group1)

group2=subset(data,data$dShock==0&data$dDeath==1&(data$tDeath-data$tShock)<=1/100)
#View(group2)# Shock time is censored but shock does not happen during a short time before the patient die, so it is likely that these patients did not die from shock.
#group1 and group2 people don't need ICD.             

group3=subset(data,data$dShock==1&(data$tDeath-data$tShock)>1/30)
#View(group3) #The patients that shock happened and live longer than 1 day after shock. (Need ICD)


censor=data[data$dShock==0&data$dDeath==0,] #For patients whose shock and death censored at almost the same time, we omit them since they may move out.
#View(censor)
censor2=censor[(censor$tDeath-censor$tShock)>=1/100,]
#View(censor2)# For patients whose shock and death censored at different time, I don't know...

library(dplyr)
other=anti_join(data, group1)
other=anti_join(other,group2)
other=anti_join(other,group3)
other=anti_join(other,censor)
View(other)#Filter above obs from data, we get other


class=rbind(group1,group2) #class contains the obs that can be classified
class=rbind(class,group3)
```

descriptive analysis
```{r}
#descriptive statistics for all classified obs
summary(class)
table(class$Sex,class$Race)
table(class$Diagnosis)

#descriptive statistics for group1 obs
summary(group1)
table(group1$Sex,group1$Race)
table(group1$Diagnosis)

#descriptive statistics for group2 obs
summary(group2)
table(group2$Sex,group2$Race)
table(group2$Diagnosis)

#descriptive statistics for group3 obs
summary(group3)
table(group3$Sex,group3$Race)
table(group3$Diagnosis)

library(ggplot2)
library(corrgram)
data=sapply(data, as.numeric)
#correlation
corrgram(data[,-c(28,29,30,31)],order = FALSE,upper.panel = panel.cor,main="correlation matrix")
cor(data[,-c(28,29,30,31)])
#cor(Diagnosis, HxMI)=0.689, cor(SysBP,DiaBP)=0.56, cor(BUN,Creatinine)=0.577, cor(QRS,LVEF)=0.536

#pca
#standardize data
dataz=scale(data[,-c(28,29,30,31)],center = T,scale = T)
dataz=as.data.frame(dataz)
pca=princomp(dataz)
summary(pca,loadings=T)
data=as.data.frame(data)

#Consider what type of people need icd most ( The patients with high shock risk)
library(survival)
#age***
##group the age
summary(data$Age)
data$Ageg[data$Age<56.0]=1
data$Ageg[56.0<=data$Age&data$Age<64.0]=2
data$Ageg[64.0<=data$Age&data$Age<72.0]=3
data$Ageg[72.0<=data$Age]=4
est_km_age=survfit(Surv(tShock, dShock) ~ strata(Ageg),data=data)
plot(est_km_age,col=c("red","blue","green","dark green"), main = "KM estimate survival function for different ages")
legend( "bottomleft",c("age1", "age2", "age3", "age4"),lty=1, col=c("red","blue","green","dark green")) 



#different sex***
est_km_sex=survfit(Surv(tShock, dShock) ~ strata(Sex),data=data)
plot(est_km_sex,col=c("red","blue"), main = "KM estimate survival function for different sex")
legend( "bottomleft",c("female", "male"),lty=1, col=c("red","blue")) 

#race***
est_km_race=survfit(Surv(tShock, dShock) ~ strata(Race),data=data)
plot(est_km_race,col=c("red","blue","green","dark green","black","orange"), main = "KM estimate survival function for different races")
legend( "bottomleft",c("AA", "AI", "AS", "NA","O","W"),lty=1, col=c("red","blue","green","dark green","black","orange")) 

#diagonosis***
est_km_diag=survfit(Surv(tShock, dShock) ~ strata(Diagnosis),data=data)
plot(est_km_diag,col=c("red","blue"), main = "KM estimate survival function for different diagnosis")
legend( "bottomleft",c("Idiopathic", "Ischemic"),lty=1, col=c("red","blue")) 

#NYHA ***
est_km_nyha=survfit(Surv(tShock, dShock) ~ strata(NYHA),data=data)
plot(est_km_nyha,col=c("red","blue","green","dark green"), main = "KM estimate survival function for different NYHA")
legend( "bottom",c("1", "2", "3", "4"),lty=1, col=c("red","blue","green","dark green")) 

#HxMI
est_km_hxmi=survfit(Surv(tShock, dShock) ~ strata(HxMI),data=data)
plot(est_km_hxmi,col=c("red","blue"), main = "KM estimate survival function for different HxMI")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#Hxdiabetes
est_km_hxdiabetes=survfit(Surv(tShock, dShock) ~ strata(HxDiabetes),data=data)
plot(est_km_hxdiabetes,col=c("red","blue"), main = "KM estimate survival function for different HxDiabetes")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#HxHTN***
est_km_hxhtn=survfit(Surv(tShock, dShock) ~ strata(HxHTN),data=data)
plot(est_km_hxhtn,col=c("red","blue"), main = "KM estimate survival function for different HxHTN")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#HxChol***
est_km_hxchol=survfit(Surv(tShock, dShock) ~ strata(HxChol),data=data)
plot(est_km_hxchol,col=c("red","blue"), main = "KM estimate survival function for different HxChol")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#BMI*** normal range (18.5,24.9)
summary(data$BMI)
data$bmig[data$BMI<18.5]=1
data$bmig[18.5<=data$BMI&data$BMI<=24.9]=2
data$bmig[24.9<data$BMI]=3
est_km_bmi=survfit(Surv(tShock, dShock) ~ strata(bmig),data=data)
plot(est_km_bmi,col=c("red","blue","green"), main = "KM estimate survival function for different bmi")
legend( "bottomleft",c("low bmi", "normal bmi", "high bmi"),lty=1, col=c("red","blue","green")) 

#Sysbp*** Normal range (90,120)
summary(data$SysBP)
data$bpg[data$SysBP<90]=1
data$bpg[90<=data$SysBP&data$SysBP<=120]=2
data$bpg[120<data$SysBP]=3
est_km_sysbp=survfit(Surv(tShock, dShock) ~ strata(bpg),data=data)
plot(est_km_sysbp,col=c("red","blue","green"), main = "KM estimate survival function for different SysBP")
legend( "bottomleft",c("low SysBP1", "normal SysBP", "high SysBP"),lty=1, col=c("red","blue","green")) 

#DiaBP*** Normal range (60,80)
#SysBP and DiaBP might be highly correlated)
data$bpg2[data$DiaBP<60]=1
data$bpg2[60<=data$DiaBP&data$DiaBP<=80]=2
data$bpg2[80<=data$DiaBP]=3
est_km_sysbp=survfit(Surv(tShock, dShock) ~ strata(bpg2),data=data)
plot(est_km_sysbp,col=c("red","blue","green"), main = "KM estimate survival function for different DiaBP")
legend( "bottomleft",c("low DiaBP", "normal DiaBP2", "high DiaBP"),lty=1, col=c("red","blue","green"))

#MedAce***
est_km_ace=survfit(Surv(tShock, dShock) ~ strata(MedAce),data=data)
plot(est_km_ace,col=c("red","blue"), main = "KM estimate survival function for different MedAce status")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#MedBeta 
est_km_beta=survfit(Surv(tShock, dShock) ~ strata(MedBeta),data=data)
plot(est_km_beta,col=c("red","blue"), main = "KM estimate survival function for different MedBeta status")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#MedNitrate***  
est_km_Nitrate =survfit(Surv(tShock, dShock) ~ strata(MedNitrate),data=data)
plot(est_km_Nitrate,col=c("red","blue"), main = "KM estimate survival function for different MedNitrate status")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#MedDigoxin ***
est_km_Digoxin =survfit(Surv(tShock, dShock) ~ strata(MedDigoxin),data=data)
plot(est_km_Digoxin,col=c("red","blue"), main = "KM estimate survival function for different MedDigoxin status")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#MedAntiarr ***
est_km_Antiarr =survfit(Surv(tShock, dShock) ~ strata(MedAntiarr),data=data)
plot(est_km_Antiarr,col=c("red","blue"), main = "KM estimate survival function for different MedAntiarr status")
legend( "bottom",c("No", "Yes"),lty=1, col=c("red","blue")) 

#BUN, normal range (7,20) ***
summary(data$BUN)
data$gbun[data$BUN<7]=1
data$gbun[7<=data$BUN&data$BUN<=20]=2
data$gbun[20<=data$BUN]=3
est_km_bun=survfit(Surv(tShock, dShock) ~ strata(gbun),data=data)
plot(est_km_sysbp,col=c("red","blue","green"), main = "KM estimate survival function for different BUN")
legend( "bottomleft",c("low BUN", "normal BUN", "high BUN"),lty=1, col=c("red","blue","green")) 

#Creatinine***, normal range (0.7,1.2)
summary(data$Creatinine)
data$gCreatinine[data$Creatinine<0.7]=1
data$gCreatinine[0.7<=data$Creatinine&data$Creatinine<=1.2]=2
data$gCreatinine[1.2<data$Creatinine]=3
est_km_Creatinine=survfit(Surv(tShock, dShock) ~ strata(gCreatinine),data=data)
plot(est_km_Creatinine,col=c("red","blue","green"), main = "KM estimate survival function for different HR")
legend( "bottomleft",c("low Creatinine", "normal Creatinine", "high Creatinine"),lty=1, col=c("red","blue","green")) 


#PR***, normal range (120,200)
summary(data$PR)
data$gpr[data$PR<120]=1
data$gpr[120<=data$PR&data$PR<=200]=2
data$gpr[200<=data$PR]=3
est_km_pr=survfit(Surv(tShock, dShock) ~ strata(gpr),data=data)
plot(est_km_pr,col=c("red","blue","green"), main = "KM estimate survival function for different pr")
legend( "bottomleft",c("low pr", "normal pr2", "high pr"),lty=1, col=c("red","blue","green")) 

#HR ,normal (60,100)
summary(data$HR)
data$ghr[data$HR<60]=1
data$ghr[60<=data$HR&data$HR<=100]=2
data$ghr[100<data$HR]=3
est_km_hr=survfit(Surv(tShock, dShock) ~ strata(ghr),data=data)
plot(est_km_pr,col=c("red","blue","green"), main = "KM estimate survival function for different HR")
legend( "bottomleft",c("low HR", "normal HR", "high HR"),lty=1, col=c("red","blue","green")) 

#***QRS,normal range (70,100), most patients have high QRS
summary(data$QRS)
data$gqrs[data$QRS<70]=1
data$gqrs[70<=data$QRS&data$QRS<=100]=2
data$gqrs[100<data$QRS]=3
est_km_qrs=survfit(Surv(tShock, dShock) ~ strata(gqrs),data=data)
plot(est_km_qrs,col=c("red","blue","green"), main = "KM estimate survival function for different QRS")
legend( "bottomleft",c("low QRS", "normal QRS", "high QRS"),lty=1, col=c("red","blue","green")) 
table(data$gqrs)

data$gqrs2[data$QRS<=100]=1
data$gqrs2[100<data$QRS]=2
est_km_qrs2=survfit(Surv(tShock, dShock) ~ strata(gqrs2),data=data)
plot(est_km_qrs2,col=c("red","blue"), main = "KM estimate survival function for different QRS")
legend( "bottomleft",c("QRS<=100", "QRS>100"),lty=1, col=c("red","blue"))


#LVEF***, normal range (55,70) , less than 35%: cardiac failure
summary(data$LVEF)
data$glvef[data$LVEF<20]=1
data$glvef[20<=data$LVEF]=2
est_km_lvef =survfit(Surv(tShock, dShock) ~ strata(LVEF),data=data)
plot(est_km_Digoxin,col=c("red","blue"), main = "KM estimate survival function for different LVEF levels")
legend( "bottom",c("LVEF<20", "LVEF>=20"),lty=1, col=c("red","blue")) 
```


##################################### Model Fitting##########################################
```{r}
attach(data)

#Estimated survival function of death
survival_death = survfit(Surv(tDeath, dDeath) ~ 1, conf.type = "log-log")
summary(survival_death)

#Estimated survival function of shock
survival_shock = survfit(Surv(tShock, dShock) ~ 1, conf.type = "log-log")
summary(survival_shock)

#Estimated hazard function of death
hazard_death = muhaz(tDeath, dDeath, bw.grid = 15, bw.method = "global", b.cor = "none")

#Estimated hazard function of shock
hazard_shock = muhaz(tShock, dShock, bw.grid = 15, bw.method = "global", b.cor = "none")

par(mfrow=c(2,2))
plot(survival_death, main = "Survival function of death", xlab = "Time", 
     ylab = "Estimated survival function of death")
plot(survival_shock, main = "Survival function of shock", xlab = "Time", 
     ylab = "Estimated survival function of shock")
plot(hazard_death, main = "Kernel-smoothed estimate hazard function of death", xlab = "Time", 
     ylab = "Estimated hazard of death")
plot(hazard_shock, main = "Kernel-smoothed estimate hazard function of shock", xlab = "Time", 
     ylab = "Estimated hazard of shock")

detach(data)
```

```{r}
#Fit the cox model of shock, full model
fit0 = coxph(Surv(tShock, dShock) ~ Age + Sex + Race + Diagnosis + NYHA + HxMI + HxDiabetes 
             + HxHTN + HxChol + BMI + SysBP + DiaBP + MedAce + MedBeta + MedNitrate + MedDiuretic
             + MedAldactone + MedDigoxin + MedAntiarr + Sodium + BUN + Creatinine + HR + PR + QRS
             + QTc + LVEF, data = data)
summary(fit0)
```

```{r}
#Based on correlation matrix and exploration, we manually select a few variables
fit1 = coxph(Surv(tShock, dShock) ~ Age + Sex + Race + Diagnosis + NYHA + HxHTN + HxChol + BMI + DiaBP 
             + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, data = data)
summary(fit1)
```

```{r}
#Stepwise variable selection based on fit0
fit2 = step(fit0, scope = list(upper = ~ Age + Sex + Race + Diagnosis + NYHA + HxMI + HxDiabetes 
                                + HxHTN + HxChol + BMI + SysBP + DiaBP + MedAce + MedBeta + MedNitrate 
                                + MedDiuretic + MedAldactone + MedDigoxin + MedAntiarr + Sodium + BUN 
                                + Creatinine + HR + PR + QRS + QTc + LVEF, 
                                lower = ~ Age))
summary(fit2)
```

```{r}
#Final selected variables
fit3 = coxph(Surv(tShock, dShock) ~ Age + Sex + Race + BMI + NYHA + HxHTN + DiaBP 
             + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, data = data)
summary(fit3)
```

```{r}
#Cox-snell residual plot
cox_snell <- function(fit) {
  cumhaz0 <- basehaz(fit) # get the baseline hazard
  Lambda0 <- approxfun(cumhaz0$time, cumhaz0$hazard, method = 'constant')
  r_C <- Lambda0(fit$y[,1]) * exp(fit$linear.predictors)
  # obtain the KM estimate of the residuals
  km_rC <- survfit(Surv(r_C, fit$y[,2]) ~ 1)
  
  df <- data.frame(x = km_rC$time, y = -log(km_rC$surv), yl = -log(km_rC$upper), 
                   yu = -log(km_rC$lower))
  df <- df[km_rC$n.risk >= 10,] # keep the estimate when at least 10 individuals are at risk
  
  figure <- ggplot(df, aes(x)) +
    geom_step(aes(y = y), colour = 'blue') + 
    geom_step(aes(y = yu), lty = 2) + geom_step(aes(y = yl), lty = 2) + 
    geom_line(aes(y = x), colour = 'red') + 
    labs(title = "Cox-Snell Residuals", x = "Residual", y = "cumulative hazard") + 
    theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
          axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
  
  return(figure)
}

cox_snell(fit3)
```

```{r}
#Fitting AFT models
#fit4 = survreg(Surv(tShock, dShock) ~ Age + Sex + Race + Diagnosis + NYHA + HxHTN + DiaBP 
#                  + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, data = #data, 
#                  dist="lognormal")
#fit5 = survreg(Surv(tShock, dShock) ~ Age + Sex + Race + Diagnosis + NYHA + HxHTN + DiaBP 
#               + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, data = data, 
#               dist="loglogistic")
fit6 = survreg(Surv(tShock, dShock) ~ Age + Sex + Race + BMI + NYHA + HxHTN + DiaBP 
               + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, data = data, 
               dist="weibull")

summary(fit6)
```


```{r}
cox_snell2 <- function(fit_survreg) {
  e_dist <- survreg.distributions[[fit_survreg$dist]]$dist # distribution of epsilon
  e_distfunc <- survreg.distributions[[e_dist]]$density # distribution functions of epsilon, the second is the survival function
  r_ST <- (log(fit_survreg$y[, 1]) - fit_survreg$linear.predictors) / fit_survreg$scale # standardized residual
  r_C <- -log(e_distfunc(r_ST)[, 2]) # Cox-Snell residual
  # obtain the KM estimate of the residuals
  km_rC <- survfit(Surv(r_C, fit_survreg$y[,2]) ~ 1)
  
  df <- data.frame(x = km_rC$time, y = -log(km_rC$surv), yl = -log(km_rC$upper), 
                   yu = -log(km_rC$lower))
  df <- df[km_rC$n.risk >= 10,] # keep the estimate when at least 10 individuals are at risk
  
  figure <- ggplot(df, aes(x)) +
    geom_step(aes(y = y), colour = 'blue') + 
    geom_step(aes(y = yu), lty = 2) + geom_step(aes(y = yl), lty = 2) + 
    geom_line(aes(y = x), colour = 'red') + 
    labs(title = paste("Checking Overall Fit of the", toupper(fit_survreg$dist), "AFT Model"), x = expression(r[Ci]), y = expression(-logS(r[Ci]))) + 
    theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
          axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
  
  return(figure)
}
cox_snell2(fit4)
cox_snell2(fit5)
cox_snell2(fit6)
```

################################## Model Assessment ############################################

```{r}
r_M_3 = residuals(fit3_3)
df = data.frame(risk_score = fit3_3$linear.predictors, status = data[-c(587, 673, 793, 844, 919),]$dShock, r_M = r_M_3)
ggplot(df, aes(x = risk_score, y = r_M_3, color = status)) +
  geom_point() + geom_smooth() + 
  labs(title = "Martingale Residuals", x = "risk score", y = "martingale residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
```
The straight line indicates no need for changepoint.

For cox model fit3, check the proportional hazard assumption
```{r}
r_SF = residuals(fit3, type = "scaledsch")
df_SF = as.data.frame(r_SF)
obstime = sort(data$tShock[data$dShock != 0])
```
```{r}
ggplot(df_SF, aes(x = obstime, y = V1)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for Age", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V2)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for Sex", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V3)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for Race", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V4)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for BMI", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V5)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for NYHA", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V6)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for HxHTN", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V7)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for DiaBP", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V8)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedAce", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V9)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedNitrate", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
```
```{r}
ggplot(df_SF, aes(x = obstime, y = V10)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedDigoxin", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
```
```{r}
ggplot(df_SF, aes(x = obstime, y = V11)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedAntiarr", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
```
```{r}
ggplot(df_SF, aes(x = obstime, y = V12)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for BUN", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V13)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for PR", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF, aes(x = obstime, y = V14)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for LVEF", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
```
The proportional hazard assumption is unsatisfied for MedDigoxin, MedAntiarr

Find reasons for nonproportionality of MedDigoxin & MedAntiarr
(1). small number of outliers (|Devience|>= 3): 587, 673, 793, 844, 919
```{r}
r_D_fit3 = residuals(fit3, type = "deviance")
df = data.frame(risk_score_fit3 = fit3$linear.predictors, r_D_fit3 = r_D_fit3)
ggplot(df, aes(x = risk_score_fit3, y = r_D_fit3)) +
  geom_point() + geom_smooth(span = 1) + 
  labs(title = "Deviance Residuals for Coxph model", x = "linear predictor", y = "deviance residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

r_D_fit3 = r_D_fit3[r_D_fit3 >= 3 | r_D_fit3 <= -3]
r_D_fit3
```

```{r}
data_2 = data[-c(587,673,793,844,919),]
fit3_2 = coxph(Surv(tShock, dShock) ~ Age + Sex + Race + Diagnosis + NYHA + HxHTN + DiaBP 
             + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, data = data_2)
r_SF_2 = residuals(fit3_2, type = "scaledsch")
df_SF_2 = as.data.frame(r_SF_2)
obstime_2 = sort(data_2$tShock[data_2$dShock != 0])

ggplot(df_SF_2, aes(x = obstime_2, y = V10)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedDigoxin", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF_2, aes(x = obstime_2, y = V11)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedAntiarr", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
```
Still significant: when sample size is large, slight departure from proportionality may be significant.

(2).consider changing nonproportional covariates into stratification factors
```{r}
fit3_3 = coxph(Surv(tShock, dShock) ~ Age + Sex + Race + Diagnosis + NYHA + HxHTN + DiaBP 
             + MedAce + MedNitrate + strata(MedDigoxin) + strata(MedAntiarr) + BUN + PR + LVEF, data = data_2)  
r_SF_3 = residuals(fit3_3, type = "scaledsch") # same procedure as above to get plots
df_SF_3 = as.data.frame(r_SF_3)
obstime_2 = sort(data_2$tShock[data_2$dShock != 0])

ggplot(df_SF_3, aes(x = obstime_2, y = V10)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedDigoxin", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))

ggplot(df_SF_3, aes(x = obstime_2, y = V11)) + 
  geom_point() + geom_smooth(span = 2) + 
  labs(title = "Scaled Schoenfeld Residuals for MedAntiarr", x = "Time (months)", y = "scaled schoenfeld residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
```
the proportional assumption holds for fit3_3
thus the final cox model is fit3_3 = coxph(Surv(tShock, dShock) ~ Age + Sex + Race + BMI + NYHA + HxHTN + DiaBP + MedAce + MedNitrate + strata(MedDigoxin) + strata(MedAntiarr) + BUN + PR + LVEF, data = data_2) 

```{r}
func1 = function(x) {pmin(x, 80)}
func2 = function(x) {pmin(x, 400)}
fit3_3 = coxph(Surv(tShock, dShock) ~ Age + Sex + Race + BMI + NYHA + HxHTN + DiaBP 
             + MedAce + MedNitrate + strata(MedDigoxin) + strata(MedAntiarr) + BUN + func1(BUN) + PR + LVEF, data = data[-c(587, 673, 793, 844, 919),])  
```
Just add func(BUN) and obtain the third cox model fit3_3
```{r}
r_M_3 = residuals(fit3_3)
df = data.frame(x = data[-c(587, 673, 793, 844, 919),]$BUN, status = data[-c(587, 673, 793, 844, 919),]$dShock, r_M = r_M_3)
ggplot(df, aes(x = data[-c(587, 673, 793, 844, 919),]$BUN, y = r_M_3, color = status)) +
  geom_point() + geom_smooth() + 
  labs(title = "Martingale Residuals", x = "BUN", y = "martingale residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
df2 = data.frame(x = data[-c(587, 673, 793, 844, 919),]$PR, status = data[-c(587, 673, 793, 844, 919),]$dShock, r_M = r_M_3)

```

```{r}

```

################################
For the AFT model:
1. check the overall fit of the weibull AFT model: previous plot about cox-snell residuals
2. deviance:
the plot shows that fit6 fits the data well
check the outliers (|Devience|>= 3): 844, 919
```{r}
r_D_fit6 = residuals(fit6, type = "deviance")
df = data.frame(risk_score_fit6 = fit6$linear.predictors, r_D_fit6 = r_D_fit6)
ggplot(df, aes(x = risk_score_fit6, y = r_D_fit6)) +
  geom_point() + geom_smooth(span = 1) + 
  labs(title = "Deviance Residuals for Weibull AFT Model", x = "linear predictor", y = "deviance residual") + 
  theme(plot.title = element_text(hjust = 0.5, size = 13), axis.title = element_text(size = 14), 
        axis.text.x = element_text(size = 11), axis.text.y = element_text(size = 12))
r_D_fit6 = r_D_fit6[r_D_fit6 >= 3 | r_D_fit6 <= -3]
r_D_fit6
outlier_2 = data[c(844,919),]
outlier_2
```
choose fit6, consider adding frailty
```{r}
data = data.frame(ID = 1:946, data)
fit6 = survreg(Surv(tShock, dShock) ~ Age + Sex + Race + BMI + NYHA + HxHTN + DiaBP 
               + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, data = data, 
               dist= "weibull")
fit6
```
Loglikelihood(fit6):-1138.2
```{r}
fit7 = parfm(Surv(tShock, dShock) ~ Age + Sex + Race + BMI + NYHA + HxHTN + DiaBP 
               + MedAce + MedNitrate + MedDigoxin + MedAntiarr + BUN + PR + LVEF, cluster = "ID", data = data, dist = "weibull", frailty = "gamma")
summary(fit7)
```

Loglikelihood(fit7): -1138.178
```{r}
LRT = 2*(-1138.178 + 1138.2)
p_value = (1 - pchisq(LRT, df = 1)) / 2
p_value
```
frailty effect is not significant. if use weibull AFT model, choose fit6
Compare coxph model and weibull proportional hazard model:

```{r}
coef_cox = fit3_3$coefficients
coef_weibull = -fit6$coefficients[2:19] / fit6$scale
data.frame(coef_cox)
data.frame(coef_weibull)
```
Compare the log-likelihood and choose the larger one / Based on coefficients above, coefficients are almost the same thus both models are reasonable.
```{r}
fit3_3$loglik
fit6$loglik
```
choose fit3_3 as the final model.
